---
title: "DAT 204"
author: "Kayla"
date: "December 8, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Final
DAT 204 Final

For the final project, I analyzed a dataset of patient information across four Kane facilities in Allegheny County. This dataset was downloaded from the Western PA Regional Data Center, which describes the dataset as "Total number of residents in each Kane Regional Center facility by race and gender. The Kane Regional Centers are skilled nursing and rehabilitation centers run by Allegheny County."

```{r}
# import librarbies for future use 
library(stringr)
library(ggplot2)

# read in csv file 
kane <- read.csv("kane.csv", stringsAsFactors=FALSE)
head(kane)
```
```{r}
pat <- (kane$patient_count) #pull out patients into vector

#  let's run some statistics on patients across all facilities 
mean <- mean(pat, na.rm = TRUE)
median <- median(pat, na.rm = TRUE)
max <- max(pat, na.rm = TRUE)
min <- min(pat, na.rm = TRUE)

print(mean)
print(median)
print(max)
print(min)

```
```{r}
# pull out number of facilities 
facilities <- unique(kane$facility)
facilities # there are four facilities in the dataframe 

# Create a dataframe for each facility by creatign a T/F vector and applying to 
# the dataframe 

fac1_tf <- str_detect(kane$facility, "1")
fac1 <- kane[fac1_tf, ]
# head(fac1)

fac2_tf <- str_detect(kane$facility, "2")
fac2 <- kane[fac2_tf, ]

fac3_tf <- str_detect(kane$facility, "3")
fac3 <- kane[fac3_tf, ]

fac4_tf <- str_detect(kane$facility, "4")
fac4 <- kane[fac4_tf, ]
```
```{r}
avg_patient <- c(mean(fac1$patient_count, na.rm = TRUE), mean(fac2$patient_count, na.rm = TRUE), mean(fac3$patient_count, na.rm = TRUE), mean(fac4$patient_count, na.rm = TRUE))

avg_patient # average patient counts for each facility

max(avg_patient) # Facility 3 had the most patients, on average

df <- data.frame("Facilities" = facilities, "Average_Patients" = avg_patient)
head(df)
```

```{r}

```

```{r}
ggplot(kane, aes(month, patient_count)) 

```

```{r}
# female_tf <- str_detect(kane$gender, 'F')
# female_df <- kane[gender_tf, ]
# head(gender_df)

facilities <- unique(kane$facility)

female_vec <- vector()

for (facility in facilities){
  fem_count <- sum(kane[kane$gender == 'F' & kane$facility == facility, 'patient_count'])
  
  female_vec <- c(female_vec, fem_count)
}

female_vec



```


```

```{r}
demo <- unique(kane$gender_race_group)
#demo


```

