---
title: "Homework 8"
author: "Kayla"
date: "November 5, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown Fun

For this homework, I am going to plot some of the data that I manipulated in the Allegheny County Dog Licenses dataset. I have pulled this dataset into a dataframe, below. 



```{r}
allegheny_county_dog_licenses <- read.csv("allegheny_county_dog_licenses.csv", stringsAsFactors=FALSE)
head(allegheny_county_dog_licenses)



```

## Rosie Dogs!

Because my family has a lovely goldendoodle named Rosie, the first thing I wanted to look at was dogs named Rosie in Allegheny County. To do so, I pulled created a true/false vector that pulled out dogs with the name Rosie, or some variation of that name. I summed it to discover that there are 121 dogs named Rosie in AC. I then put this into a separate dataframe. 


```{r echo=FALSE}

library(stringr)

rosie_tf <- str_detect(allegheny_county_dog_licenses$DogName, "ROSIE") #pull out dogs named Rosie
sum(rosie_tf) # there are 121, hooray!
rosie_df <- allegheny_county_dog_licenses[rosie_tf, ] # create df
head(rosie_df)

```

## Goldendoodles! 

The next thing I did was pull out dogs named Rosie that are goldendoodles. To do so, I made a true/false vector to sort for the breed "GOLDENDOODLE" and applied it to the Rosie dataframe. I then summed it to discover there are 8 goldendoodles in AC. 
```{r}

doodle_rosie_tf <- str_detect(rosie_df$Breed, "GOLDENDOODLE") # from dogs named Rosie, pull out goldendoodles 
sum(doodle_rosie_tf) # there are 8, double hooray! 
doodle_rosie_df <- rosie_df[doodle_rosie_tf, ] # create df 
head(doodle_rosie_df)

```

## Let's plot! 

Next, I decided to plot my 8 happy doods named Rosie. 
```{r}
library('ggplot2')

ggplot(data = doodle_rosie_df, aes(x=Color))+
  geom_bar(color="black", fill="lightgoldenrod2")
```



## Rosalies! (Or lack thereof ...)
My Rosie's full name is Rosalie, but she lives in Washington County with my parents. I decided to see if she has an Allegheny County counterpart. First, I created a true/false vector to check the original dataset for dogs named Rosalie. 
```{r}
rosalie_tf <- str_detect(allegheny_county_dog_licenses$DogName, "ROSALIE") #pull out dogs named Rosalie 
sum(rosalie_tf)
```
 There are no dogs in Allegheny County names Rosalie! My Rosie is [truly unique](https://www.instagram.com/rosaliethedoodle/).
 
 
## Lhasa Apso Data!
Next, I decided to look at data for dogs that are a Lhasa Apso mix. My family had a loyal little Lhasa Apso Poodle mix named [Oreo](https://www.instagram.com/p/B9mC0uvhBk5/) for 15 years. 

```{r}
# Pull out all the Lhasa Apso mixes
lhasa_mix_tf <- str_detect(allegheny_county_dog_licenses$Breed, "LHASA APSO MIX")
sum(lhasa_mix_tf) # there are 21 Lhasa Apso mixes
lhasa_mix_df <- allegheny_county_dog_licenses[lhasa_mix_tf,]
```
## Fun with Names! 

With such a small dataframe, let's look at their names 
```{r}
lhasa_mix_names <- unique(lhasa_mix_df$DogName) # pull out unique names into a vector
lhasa_mix_names
length(lhasa_mix_names)
```
There are 21 Lhasa Apso mixes in Allegheny County and when we run unique on the DogName column, we get 21 names. This means each Lhasa Apso mix in Allegheny County has a unique name. While this is cute, we can't plot much with this. So, let's look at something else! Instead, I will compare Lhasa Apsos and Lhasa Apso mixes. 

```{r}
# pull out all dog breeds with "Lhasa" in the breed name 
all_lhasas_tf <- str_detect(allegheny_county_dog_licenses$Breed, "LHASA")
sum(all_lhasas_tf)

# pull all Lhasas into a dataframe 
all_lhasas_df <- allegheny_county_dog_licenses[all_lhasas_tf,]
all_lhasas_df


```
## Let's plot!
Now we have both purebred Lhasa Apsos and Lhasa Apso mixes in a data frame. 


```{r}

ggplot(data = all_lhasas_df, aes(x=Breed))+
  geom_bar(color="black", fill="mediumorchid2")+
  ggtitle("Lhasa Apsos - Purebreds vs Mixes")+
  ylab("Total")

```


Obviously, we have a lot more purebred Lhasa Apso dogs than mixes in Allegheny County! 
